---
title: "1-week VaR estimation: DAX & SPX"
link-citations: yes
output:
  pdf_document:
    toc: yes
    toc_depth: 3
    latex_engine: xelatex
  html_notebook:
    toc: yes
    toc_depth: 3
    toc_float: yes
  html_document:
    toc: yes
    toc_depth: 3
    df_print: paged
  word_document:
    toc: yes
    toc_depth: '3'
bibliography: references.bib
header-includes: \usepackage{booktabs}
editor_options:
  markdown:
    wrap: 72
---
## Individual contributions
22085374, 23006530, 22008197: Report writing and synthesis

\noindent 23144606, 23097365, 23122857: Code & calculations


\newpage

# 1. Introduction

\noindent Classical $VaR$ often relies on joint normality and linear
correlation, which can be unreliable for financial returns due to
skewness, heavy tails, volatility clustering, and asymmetric dependence
in extremes. Copulas offer a principled alternative: they model the
dependence structure separately from the marginal distributions,
allowing more realistic tail behaviour and flexible joint modelling.

\noindent A key requirement for static copula estimation is that the
copula inputs are approximately i.i.d. over time. Concretely, after
transforming each marginal to uniforms, the pseudo-observations
$(u_{\mathrm{SPX},t},u_{\mathrm{DAX},t})$ should behave like independent
draws from a time-invariant bivariate distribution. Raw equity returns
often violate this because volatility clustering and other dynamics
induce serial dependence in the conditional distribution.

\noindent Our workflow is therefore: (i) exploratory diagnostics to
check for heavy tails, skewness, autocorrelation and volatility
clustering; (ii) fit ARMA-GARCH models to each index to capture
conditional mean dynamics and conditional heteroskedasticity, yielding
standardised innovations that are closer to i.i.d.; (iii) apply the
probability integral transform (PIT) to these innovations to obtain
uniform pseudo-observations; (iv) fit and select a copula on
$(u_{\mathrm{SPX},t},u_{\mathrm{DAX},t})$; and (v) estimate 1-week $95$%
and $99$% $VaR$ for the weekly rebalanced portfolio via Monte Carlo
simulation under the fitted marginal--copula model.

# 2. Main Task

## 2.1 Initialisation

```{r, message=FALSE}
library(tseries)
library(fGarch)
library(KScorrect)
library(VineCopula)
library(goftest)
library(rugarch)


# Data Importing and Preparation
spx_w <- read.csv("^spx_w.csv")
dax_w <- read.csv("^dax_w.csv")


# Extract close prices
spx_close<-spx_w$Close 
dax_close<-dax_w$Close


# Convert Date column
spx_w$Date <- as.Date(spx_w$Date)
dax_w$Date <- as.Date(dax_w$Date)

# Keep only Date + Close
spx_close_df <- spx_w[, c("Date", "Close")]
dax_close_df <- dax_w[, c("Date", "Close")]

# Set start and end date
START_DATE <- as.Date("2000-01-02")
END_DATE   <- as.Date("2025-12-28")
dax_close_df <- dax_close_df[dax_close_df$Date >= START_DATE & dax_close_df$Date <= END_DATE, ]
spx_close_df <- spx_close_df[spx_close_df$Date >= START_DATE & spx_close_df$Date <= END_DATE, ]

# Rename close columns so they don't collide
names(spx_close_df)[2] <- "SPX"
names(dax_close_df)[2] <- "DAX"


prices <- merge(spx_close_df, dax_close_df, by = "Date", all = FALSE)
prices <- prices[complete.cases(prices), ]     # Drop any remaining NAs
prices <- prices[order(prices$Date), ]         # Sort by date
```

## 2.2 Log-returns and portfolio definition

```{r}
# Creation of the log-returns of the indices
rSPX <- diff(log(prices$SPX))
rDAX <- diff(log(prices$DAX))

returns <- data.frame(
  Date = prices$Date[-1],
  rSPX = rSPX,
  rDAX = rDAX
)

# equally weighted portfolio (log-return approximation)
returns$rP <- 0.5 * returns$rSPX + 0.5 * returns$rDAX

# equally weighted portfolio with weekly rebalancing (consistent with your simulation definition)
R1 <- exp(returns$rSPX) - 1
R2 <- exp(returns$rDAX) - 1
returns$rP_rebal <- log(1 + 0.5*R1 + 0.5*R2)
```

Pricing data was obtained from Stooq [@stooq_spx_nodate]. Let $p_{j,t}$
be the weekly closing price for index j
$\in \{\mathrm{SPX},\, \mathrm{DAX}\}$ at week $t$. Weekly log-return:

$$
r_{j,t} \;=\; \log(P_{j,t}) - \log(P_{j,t-1})
\;=\; \log\!\left(\frac{P_{j,t}}{P_{j,t-1}}\right)
$$ To construct an equally weighted Weekly portfolio log-return:

$$
R_{p,t} = \frac{1}{2}\,(e^{r_{\mathrm{SPX},t}}-1) + \frac{1}{2}\,(e^{r_{\mathrm{DAX},t}} - 1),
\qquad
r_{p,t} = \log\!\left(1 + R_{p,t}\right).
$$

\noindent We define the portfolio loss fraction for 1-week (relative to
the current value $V_0$) as

$$
L_t \;=\; -\,r_{p,t}.
$$

\noindent Thus, $\mathrm{VaR}_c$ at confidence level
$c \in \{0.95, 0.99\}$ is the $c$-quantile of $L$:

$$
\mathrm{VaR}_c \;=\; \mathrm{Quantile}_c(L).
$$

## 2.3 Exploratory analysis

```{r levels_returns,fig.width=10, fig.height=5, echo=FALSE, fig.cap="Weekly normalised index levels for the SPX and DAX (left) and their weekly log returns (right) from 2000–2025."}

par(mfrow = c(1, 2))
# Normalised index levels
prices$SPX_100 <- 100 * prices$SPX / prices$SPX[1]
prices$DAX_100 <- 100 * prices$DAX / prices$DAX[1]

plot(prices$Date, prices$SPX_100, type="l", col="green", xlab="Date", ylab="Index (start=100)", main="Weekly index levels (normalised)")
lines(prices$Date, prices$DAX_100, col="firebrick")

legend("topleft",legend=c("SPX (100=start)","DAX (100=start)"), col=c("green","firebrick"), lty=1, bty="n")

# Weekly log returns
plot(returns$Date, returns$rSPX, type="l", col="green", xlab="Date", ylab="Weekly log returns", main="Weekly log returns")
lines(returns$Date, returns$rDAX, col="firebrick")
abline(h=0, lty=2, col="gray60")

legend("topright",
       legend=c("SPX","DAX"),
       col=c("green","firebrick"), lty=1, bty="n")
```

To compare long-run behaviour on a common scale, we normalise each index
to start at 100 and plot weekly index levels (Figure 1). Both SPX and
DAX trend upward over the sample with pronounced drawdowns during stress
periods (notably 2008--2009 and 2020). The corresponding weekly
log-return series fluctuates around zero and shows clear volatility
clustering (bursts of large moves followed by calmer periods),
motivating a conditional heteroskedastic model in each marginal series.

```{r hist_qq, fig.width=10, fig.height=8, echo=FALSE, fig.cap= "Histograms and normal Q–Q plots of weekly log-returns for the SPX and DAX."}
par(mfrow = c(2, 2))
hist(returns$rSPX, main="SPX weekly log-returns", xlab="")
qqnorm(returns$rSPX); qqline(returns$rSPX)
hist(returns$rDAX, main="DAX weekly log-returns", xlab="")
qqnorm(returns$rDAX); qqline(returns$rDAX)

```

```{r, echo=FALSE, results='hide'}
desc <- function(x){
	x <- na.omit(x)
	m <- mean(x); s <- sd(x)
	skew <- mean((x-m)^3)/s^3
	kurt <- mean((x-m)^4)/s^4 - 3
	jb <- jarque.bera.test(x)
	
	c(mean=m, sd=s, skew=skew, ex_kurt=kurt, JB = unname(jb$statistic), JB_p=jb$p.value)
}

rbind(SPX=desc(returns$rSPX), DAX=desc(returns$rDAX))
```

\begin{table}[h!]
\centering
\begin{tabular}{lcccccc}
\toprule
Index & Mean & SD & Skew & Excess Kurtosis & JB & JB\_p \\
\midrule
SPX & 0.0011438784 & 0.02483206 & -0.8709850 & 7.071785 & 3009.296 & 0 \\
DAX & 0.000923694 & 0.03083369 & -0.8803177 & 6.782680 & 2785.880 & 0 \\
\bottomrule
\end{tabular}
\caption{Summary statistics and Jarque-Bera (JB) tests for weekly log-returns of SPX and DAX.}
\end{table}

Histograms and normal Q--Q plots (Figure 2) show strong departures from
normality in both tails, with particularly large deviations in the lower
tail. This visual evidence is consistent with the numerical summaries:
both return series have pronounced negative skewness (approximately
$-0.87$ to $-0.88$) and substantial excess kurtosis (approximately $6.8$
to $7.1$), indicating asymmetric heavy lower tails where extreme
negative returns occur more frequently and with larger magnitude than
extreme positive returns.

Normality is assessed formally using the Jarque--Bera (JB) where small
$p$-values provide evidence against normality. The JB test strongly
rejects normality for both indices (SPX: JB $\approx 3009$; DAX: JB
$\approx 2786$; both with $p<2.2\times 10^{-16}$) (Table 1). Overall,
the plots and tests motivate heavy-tailed, potentially skewed marginal
innovations and dependence modelling beyond using simple linear
correlation.

```{r acf_pacf_log, fig.width=10, fig.height=10, echo=FALSE, fig.cap= "ACF and PACF plots for SPX and DAX weekly log-returns and squared returns."}

par(mfrow = c(3, 2))
acf(returns$rSPX, main="ACF: SPX returns")
acf(returns$rDAX, main="ACF: DAX returns")
acf(returns$rSPX^2, main="ACF: SPX squared returns")
acf(returns$rDAX^2, main="ACF: DAX squared returns")
pacf(returns$rSPX, main="PACF: SPX returns")
pacf(returns$rDAX, main="PACF: DAX returns")
```

The ACF/PACF plots (Figure 3) show that most return autocorrelations lie
within the 95% bounds, so any mean dependence is weak and, if present,
likely low order. By contrast, the ACF of squared returns is clearly
positive at short lags for both indices, indicating volatility
clustering and motivating a GARCH-type variance model. Combined with the
large excess kurtosis (Table 1), this points to an ARMA--GARCH
specification with low-order mean dynamics and heavy-tailed innovations;
we assess the ARMA order formally in the next section.

## 2.4 Conditional mean model selection

Following the Box--Jenkins approach, we compare low-order AR(p)
specifications with p=0,1,2,3 as return ACF/PACF indicated at most weak
short-lag dependence and no stable structure consistent with
higher-order mean dynamics:

$$
r_{j,t} = \mu_{j,t} + \varepsilon_{j,t},
\qquad
\mu_{j,t} = c_j + \sum_{i=1}^{p}\phi_{j,i}\,r_{j,t-i} + \sum_{k=1}^{q}\theta_{j,k}\,\varepsilon_{j,t-k}.
$$

Candidate models are estimated via maximum likelihood and compared using
AIC/BIC (prioritising BIC to favour parsimony). The selected
specification is then checked using Ljung--Box tests on the standardised
residuals, which test the joint null hypothesis of no autocorrelation up
to a chosen lag (so rejection would indicate remaining serial dependence
and an inadequate mean specification).

\begin{table}[h!]
\centering
\begin{tabular}{ccc}
\toprule
$p$ & AIC & BIC \\
\midrule
0 & -6171.360 & -6160.935 \\
1 & -6177.608 & -6161.971 \\
2 & -6178.569 & -6157.720 \\
3 & -6178.859 & -6152.798 \\
\bottomrule
\end{tabular}
\caption{AIC and BIC values for AR($p$) models fitted to SPX returns.}
\end{table}

```{r, echo=FALSE, results='hide'}
# SPX: AR(0) to AR(3)
spx0 <- arima(returns$rSPX, order=c(0,0,0), include.mean=TRUE)
spx1 <- arima(returns$rSPX, order=c(1,0,0), include.mean=TRUE)
spx2 <- arima(returns$rSPX, order=c(2,0,0), include.mean=TRUE)
spx3 <- arima(returns$rSPX, order=c(3,0,0), include.mean=TRUE)

cbind(
  p   = 0:3,
  AIC = c(AIC(spx0), AIC(spx1), AIC(spx2), AIC(spx3)),
  BIC = c(BIC(spx0), BIC(spx1), BIC(spx2), BIC(spx3))
)
```

```{r, echo=FALSE, results='hide' }
fit_ar1_spx <- arima(returns$rSPX, order = c(1,0,0), include.mean = TRUE)
Box.test(residuals(fit_ar1_spx), lag = 10, type = "Ljung-Box")
```

\begin{table}[htbp]
\centering
\label{tab:lb_dax_mean}
\begin{tabular}{lccc}
\toprule
Model & Q statistic & df & p-value \\
\midrule
ARMA(1,0) & 16.261 & 10 & 0.09239 \\
\bottomrule
\end{tabular}
\caption{Ljung--Box test on fitted standardised residuals for selected SPX mean model.}
\end{table}

For $SPX$, BIC is minimised at $p=1$ (Table 2), and the LB tests on
fitted residuals across 10 lags indicated no remaining linear
autocorrelation at standard lags (p\>0.05) (Table 3). Therefore we
proceed with an ARMA(1,0) mean specification.

\begin{table}[h!]
\centering
\begin{tabular}{ccc}
\toprule
$p$ & AIC & BIC \\
\midrule
0 & -5584.287 & -5573.862 \\
1 & -5583.067 & -5567.430 \\
2 & -5584.443 & -5563.594 \\
3 & -5593.600 & -5567.538 \\
\bottomrule
\end{tabular}
\caption{AIC and BIC values for AR($p$) models fitted to DAX returns.}
\end{table}

```{r, echo=FALSE, results='hide'}
# DAX: AR(0) to AR(3)
dax0 <- arima(returns$rDAX, order=c(0,0,0), include.mean=TRUE)
dax1 <- arima(returns$rDAX, order=c(1,0,0), include.mean=TRUE)
dax2 <- arima(returns$rDAX, order=c(2,0,0), include.mean=TRUE)
dax3 <- arima(returns$rDAX, order=c(3,0,0), include.mean=TRUE)

cbind(
  p   = 0:3,
  AIC = c(AIC(dax0), AIC(dax1), AIC(dax2), AIC(dax3)),
  BIC = c(BIC(dax0), BIC(dax1), BIC(dax2), BIC(dax3))
)
```

```{r, echo=FALSE, results='hide'}
fit_ar0_dax <- arima(returns$rDAX, order = c(0,0,0), include.mean = TRUE)
Box.test(residuals(fit_ar0_dax), lag = 10, type = "Ljung-Box")
fit_ar1_dax <- arima(returns$rDAX, order = c(1,0,0), include.mean = TRUE)
Box.test(residuals(fit_ar1_dax), lag = 10, type = "Ljung-Box")
fit_ar2_dax <- arima(returns$rDAX, order = c(2,0,0), include.mean = TRUE)
Box.test(residuals(fit_ar2_dax), lag = 10, type = "Ljung-Box")
fit_ar3_dax <- arima(returns$rDAX, order = c(3,0,0), include.mean = TRUE)
Box.test(residuals(fit_ar3_dax), lag = 10, type = "Ljung-Box")
```

\begin{table}[htbp]
\centering
\label{tab:lb_dax_mean}
\begin{tabular}{lccc}
\toprule
Model & Q statistic & df & p-value \\
\midrule
ARMA(0,0) & 24.5770 & 10 & 0.006207 \\
ARMA(1,0) & 22.7740 & 10 & 0.01161 \\
ARMA(2,0) & 19.2640 & 10 & 0.03704 \\
ARMA(3,0) & 7.7231 & 10 & 0.6559 \\
\bottomrule
\end{tabular}
\caption{Ljung--Box tests on fitted standardised residuals for DAX mean model candidates.}
\end{table}

\noindent For $DAX$, the BIC-selected ARMA$(0,0)$ specification (Table
4) produced a significant Ljung--Box statistic on the fitted
(standardised) residuals (p-value $<0.05$) (Table 5), indicating
remaining serial dependence in the conditional mean. We therefore
evaluated higher-order mean specifications, and retained the ARMA$(3,0)$
model as the lowest-order choice for which the Ljung--Box test failed to
reject the null of no autocorrelation at the 5% level (p-value $>0.05$)
(Table 5).

\noindent Therefore the final conditional mean models are ARMA(1,0) for
$SPX$ and ARMA(3,0) for $DAX$. \newpage

## 2.5 Conditional variance model selection

```{r, echo=FALSE, results='hide'}
Box.test(residuals(fit_ar1_spx)^2, lag = 10, type = "Ljung-Box")
Box.test(residuals(fit_ar3_dax)^2, lag = 10, type = "Ljung-Box")
```

\begin{table}[htbp]
\centering
\label{tab:lb_sqres}
\begin{tabular}{lccc}
\toprule
Series / fitted model & Q statistic & df & p-value \\
\midrule
SPX (mean model residuals)$^2$ & 330.31 & 10 & $< 2.2\times10^{-16}$ \\
DAX (mean model residuals)$^2$ & 267.80 & 10 & $< 2.2\times10^{-16}$ \\
\bottomrule
\end{tabular}
\caption{Ljung--Box tests on squared residuals (testing for remaining ARCH clustering).}
\end{table}

The Ljung--Box test applied to the squared residuals from the fitted
ARMA mean models is significant for both series (p-value $<0.05$),
corroborating the positive autocorrelation observed in the ACF of
squared mean-model residuals (Table 6). This provides formal evidence of
ARCH-type dependence (volatility clustering) and motivates a GARCH
specification for the conditional variance:

$$
\varepsilon_{j,t}=\sigma_{j,t}z_{j,t}, \qquad 
\sigma_{j,t}^{2}=\omega_j+\alpha_j\varepsilon_{j,t-1}^{2}+\beta_j\sigma_{j,t-1}^{2},
$$ where $z_{j,t}$ are i.i.d. innovations.

Holding the conditional mean fixed at the selected ARMA orders (SPX:
ARMA$(1,0)$; DAX: ARMA$(3,0)$), we compare low-order GARCH$(p,q)$
specifications with $(p,q)\in\{(1,1),(1,2),(2,1),(2,2)\}$ via maximum
likelihood. Since both return series exhibit heavy tails, these
candidate GARCH models are initially estimated under Student-$t$
innovations to avoid variance underestimation under Gaussian errors.
Models are compared using AIC/BIC (prioritising BIC to favour
parsimony).

\begin{table}[h!]
\centering
\begin{tabular}{lcccc}
\toprule
Model & AIC & BIC & SIC & HQIC \\
\midrule
g11 & -4.909832 & -4.886769 & -4.909871 & -4.901197 \\
g12 & -4.908842 & -4.881935 & -4.908895 & -4.898768 \\
g21 & -4.908750 & -4.881843 & -4.908803 & -4.898675 \\
g22 & -4.907374 & -4.876623 & -4.907443 & -4.895860 \\
\bottomrule
\end{tabular}
\caption{Information criteria for candidate GARCH models fitted to SPX returns.}
\label{tab:garch_spx}
\end{table}
\begin{table}[h!]
\centering
\begin{tabular}{lcccc}
\toprule
Model & AIC & BIC & SIC & HQIC \\
\midrule
g11 & -4.421803 & -4.391052 & -4.421872 & -4.410289 \\
g12 & -4.420576 & -4.385981 & -4.420664 & -4.407623 \\
g21 & -4.420293 & -4.385698 & -4.420380 & -4.407340 \\
g22 & -4.419101 & -4.380663 & -4.419209 & -4.404709 \\
\bottomrule
\end{tabular}
\caption{Information criteria for candidate GARCH models fitted to DAX returns.}
\label{tab:garch_dax}
\end{table}

```{r, echo=FALSE, results='hide'}
# Fit candidate SPX variance models (keep mean fixed at AR(1))

spx_g11 <- garchFit(~ arma(1,0) + garch(1,1), data=returns$rSPX, trace=FALSE, cond.dist="std")
spx_g12 <- garchFit(~ arma(1,0) + garch(1,2), data=returns$rSPX, trace=FALSE, cond.dist="std")
spx_g21 <- garchFit(~ arma(1,0) + garch(2,1), data=returns$rSPX, trace=FALSE, cond.dist="std")
spx_g22 <- garchFit(~ arma(1,0) + garch(2,2), data=returns$rSPX, trace=FALSE, cond.dist="std")

# Compare info criteria
spx_ic <- rbind(
  g11 = spx_g11@fit$ics,
  g12 = spx_g12@fit$ics,
  g21 = spx_g21@fit$ics,
  g22 = spx_g22@fit$ics
)
spx_ic
```

```{r, echo=FALSE, warning=FALSE, results='hide'}
### DAX: AR(3) + choose GARCH(p,q)

# Fit candidate variance models (keep mean fixed at AR(3))
dax_g11 <- garchFit(~ arma(3,0) + garch(1,1), data=returns$rDAX, trace=FALSE, cond.dist="std")
dax_g12 <- garchFit(~ arma(3,0) + garch(1,2), data=returns$rDAX, trace=FALSE, cond.dist="std")
dax_g21 <- garchFit(~ arma(3,0) + garch(2,1), data=returns$rDAX, trace=FALSE, cond.dist="std")
dax_g22 <- garchFit(~ arma(3,0) + garch(2,2), data=returns$rDAX, trace=FALSE, cond.dist="std")

# Compare info criteria
dax_ic <- rbind(
  g11 = dax_g11@fit$ics,
  g12 = dax_g12@fit$ics,
  g21 = dax_g21@fit$ics,
  g22 = dax_g22@fit$ics
)
dax_ic
```

For DAX, the information criteria agree: GARCH$(1,1)$ attains the lowest
AIC and BIC among the candidate orders, so we retain GARCH$(1,1)$ for
the conditional variance (Table 8).

For SPX, the information criteria are very close across GARCH$(1,1)$,
GARCH$(1,2)$, GARCH$(2,1)$ and GARCH$(2,2)$: AIC is minimised by
GARCH$(1,1)$, while BIC is marginally lower for GARCH$(2,2)$ (Table 7).
Given the
negligible differences and the preference for a parsimonious volatility
specification, we select GARCH$(1,1)$ for SPX as well.

Conditional on the selected ARMA--GARCH orders, we now compare
alternative innovation distributions within this structure and select
the final conditional distribution using likelihood-based criteria
(Table 9 and Table 10).

\begin{table}[h!]
\centering
\begin{tabular}{lcccc}
\toprule
Distribution & AIC & BIC\\
\midrule
Normal & -4.854768 & -4.835549\\
Student-t & -4.909832 & -4.886769\\
Skewed Student-t & -4.941944 & -4.915037\\
Skewed Normal & -4.910392 & -4.887328\\
Skewed Generalised Error Distribution & -4.935027 & -4.908120\\
Generalised Error Distribution & -4.896581 & -4.873517\\
\bottomrule
\end{tabular}
\caption{Information criteria for alternative innovation distributions in the ARMA–GARCH model for SPX returns.}
\label{tab:innov_spx}
\end{table}
\begin{table}[h!]
\centering
\begin{tabular}{lcccc}
\toprule
Distribution & AIC & BIC\\
\midrule
Normal & -4.354826 & -4.327918\\
Student-t & -4.421803 & -4.391052\\
Skewed Student-t & -4.459454 & -4.424859\\
Skewed Normal & -4.417081 & -4.386330\\
Skewed Generalised Error Distribution & -4.447922 & -4.413328\\
Generalised Error Distribution & -4.402475 & -4.371724\\
\bottomrule
\end{tabular}
\caption{Information criteria for alternative innovation distributions in the ARMA–GARCH model for DAX returns.}
\label{tab:innov_dax}
\end{table}

```{r, echo=FALSE, results='hide'}
# SPX
model11=garchFit(formula=~arma(1,0)+garch(1,1),data=returns$rSPX,trace=F,cond.dist="norm")
model12=garchFit(formula=~arma(1,0)+garch(1,1),data=returns$rSPX,trace=F,cond.dist="std")
model13=garchFit(formula=~arma(1,0)+garch(1,1),data=returns$rSPX,trace=F,cond.dist="sstd")
model14=garchFit(formula=~arma(1,0)+garch(1,1),data=returns$rSPX,trace=F,cond.dist="snorm")
model15=garchFit(formula=~arma(1,0)+garch(1,1),data=returns$rSPX,trace=F,cond.dist="sged")
model16=garchFit(formula=~arma(1,0)+garch(1,1),data=returns$rSPX,trace=F,cond.dist="ged")
ic=rbind(model11@fit$ics,model12@fit$ics,model13@fit$ics,model14@fit$ics,
model15@fit$ics,model16@fit$ics)
ic
```

```{r, echo=FALSE, results='hide'}
# DAX
# DAX
model21=garchFit(formula=~arma(3,0)+garch(1,1),data=returns$rDAX,trace=F,cond.dist="norm")
model22=garchFit(formula=~arma(3,0)+garch(1,1),data=returns$rDAX,trace=F,cond.dist="std")
model23=garchFit(formula=~arma(3,0)+garch(1,1),data=returns$rDAX,trace=F,cond.dist="sstd")
model24=garchFit(formula=~arma(3,0)+garch(1,1),data=returns$rDAX,trace=F,cond.dist="snorm")
model25=garchFit(formula=~arma(3,0)+garch(1,1),data=returns$rDAX,trace=F,cond.dist="sged")
model26=garchFit(formula=~arma(3,0)+garch(1,1),data=returns$rDAX,trace=F,cond.dist="ged")
ic2=rbind(model21@fit$ics,model22@fit$ics,model23@fit$ics,model24@fit$ics,
model25@fit$ics,model26@fit$ics)
ic2
```

Across both indices, the skewed Student-$t$ innovation attains the
lowest information criteria, so it is retained as the best-fitting
conditional distribution (Table 9 and Table 10). This aligns with the
sample summary statistics (excess kurtosis and mild skewness), implying
heavy-tailed and asymmetric return shocks that are not captured by
Gaussian innovations.

## 2.6 ARMA-GARCH model diagnostics

The adequacy of the \emph{final} ARMA--GARCH specification is assessed
using Ljung--Box tests on the standardised residuals and on their
squares (Table 11).

\begin{table}[htbp]
\centering
\label{tab:lb_squared_residuals}
\begin{tabular}{lccc}
\toprule
Index & Q statistic & df & p-value \\
\midrule
SPX & 9.6667 & 10 & 0.4702 \\
DAX & 10.564 & 10 & 0.3924 \\
\bottomrule
\end{tabular}
\caption{Ljung--Box tests on squared standardised residuals of the final ARMA--GARCH models.}
\end{table}

```{r, echo=FALSE, results='hide'}
Box.test(residuals(model13, standardize= T)^2, lag = 10, type = "Ljung-Box")
Box.test(residuals(model23, standardize= T)^2, lag = 10, type = "Ljung-Box")
Box.test(residuals(model13, standardize= T), lag = 10, type = "Ljung-Box")
Box.test(residuals(model23, standardize= T), lag = 10, type = "Ljung-Box")
```

```{r, fig.width=10, fig.height=8, echo=FALSE, fig.cap= "ACF of standardised residuals and standardised squared residuals for the final ARMA–GARCH models fitted to SPX and DAX returns."}


# standardised residuals
z_spx <- residuals(model13, standardize=TRUE)
z_dax <- residuals(model23, standardize=TRUE)

par(mfrow=c(2,2))
acf(z_spx, main="ACF: SPX std residuals")
acf(z_spx^2, main="ACF: SPX squared std residuals")
acf(z_dax, main="ACF: DAX std residuals")
acf(z_dax^2, main="ACF: DAX squared std residuals")
```

```{r, fig.width=10, fig.height=10, echo=FALSE, fig.cap= "Time series of standardised residuals from the selected ARMA–GARCH models for SPX and DAX."}

par(mfrow=c(2,1))
plot(z_spx, type = "l", main = "SPX: standardised residuals", xlab = "t", ylab = "z_t")
abline(h = 0, lty = 2); grid()

plot(z_dax, type = "l", main = "DAX: standardised residuals", xlab = "t", ylab = "z_t")
abline(h = 0, lty = 2); grid()

```

\newpage
Although the ACFs of the standardised residuals in Figure 4 indicate
minor residual autocorrelation at a few lags, there is no strong
systematic pattern, suggesting that the ARMA component has captured most
of the dependence in the conditional mean, in line with Tables 3 and 5.
Likewise, the ACFs of the squared standardised residuals in Figure 4
show only limited remaining dependence, while the Ljung-Box results for
the final model in Table 11 indicate no material residual
autocorrelation in the squared residuals. Consistent with this, the time
series of standardised residuals in Figure 5 are centred around zero and
show no obvious remaining volatility clustering, indicating that the
ARMA-GARCH specifications have largely filtered the conditional mean and
variance dynamics. The resulting innovations are therefore approximately
i.i.d., making them suitable for the probability integral transform. On
this basis, we retain the ARMA-GARCH specification as the final model
and proceed with the remaining diagnostic checks.

## 2.7 Probability integral transform & copula selection

Let $\hat{z}_{j,t}$ denote the fitted standardised residuals obtained
from the fitted ARMA--GARCH marginal model for each series $j$ , defined
as:

$$
\hat{\varepsilon}_{j,t} = r_{j,t} - \hat{\mu}_{j,t},
\qquad
\hat{z}_{j,t} = \frac{\hat{\varepsilon}_{j,t}}{\hat{\sigma}_{j,t}}.
$$

Since filtering is applied marginally, it removes serial dependence in
the conditional mean and variance of each series independently, but does
not eliminate cross-sectional dependence.\
The PIT is applied under the fitted skewed students-$t$ (sstd)
distribution to map the standardised residuals to the unit interval:

$$
u_{j,t} = \hat{F}_{\mathrm{sstd}}\!\left(\hat{z}_{j,t};\,\hat{\nu}_j, \hat{\xi}_j\right),
$$

where $\hat{\nu_{j}}$ and $\hat{\xi_{j}}$ are the fitted degrees of
freedom and skewness parameters for series $j$. Under correct marginal
specification, $u_{j,t} \sim \text{i.i.d. } U(0,1)$. To avoid degeneracy
at the boundary, the pseudo-observations are clamped to
$[10^{-6},\,1-10^{-6}]$.

```{r}
#extract sstd condition/innovation
parspx <- model13@fit$par
pardax <- model23@fit$par

#extract the parameters from the distribution
nuspx <- unname(parspx["shape"])
xispx <- unname(parspx["skew"])
nudax <- unname(pardax["shape"])
xidax <- unname(pardax["skew"])
#apply the distribution to the standardized residuals 
uspx <- psstd(z_spx, nu = nuspx, xi = xispx)
udax <- psstd(z_dax, nu = nudax, xi = xidax)

#clamp
eps <- 1e-6
uspx <- pmin(pmax(uspx, eps), 1 - eps)
udax <- pmin(pmax(udax, eps), 1 - eps)

#combine pairs to fit a copula
U <- na.omit(cbind(uspx, udax))
colnames(U) <- c("USPX","UDAX")
```

```{r, fig.width=10, fig.height=5, echo=FALSE, fig.cap= "Pseudo-observations for SPX and DAX generated from the fitted skewed Student-t innovation distributions. The histograms are approximately uniform, therefore supporting their suitability for copula modelling."}
par(mfrow = c(1, 2))
hist(uspx)
hist(udax)
```

Uniformity of the pseudo-observations is assessed via the
Lilliefors-corrected Kolmogorov--Smirnov (KS) and Anderson--Darling
tests.

\begin{table}[h!]
\centering
\begin{tabular}{lcc}
\toprule
Series & KS p-value & AD p-value \\
\midrule
SPX & 0.024 & 0.01216645 \\
DAX & 0.3472 &  0.1760574 \\
\bottomrule
\end{tabular}
\caption{Kolmogorov--Smirnov and Anderson--Darling test p-values for pseudo-observations.}
\end{table}

```{r, echo=FALSE, results='hide'}
KStest1 <- LcKS(uspx, cdf = "punif")
KStest1$p.value
ADtest1 <- ad.test(uspx, null = "punif")
ADtest1$p.value

KStest2 <- LcKS(udax, cdf = "punif")
KStest2$p.value
ADtest2 <- ad.test(udax, null = "punif")
ADtest2$p.value

```

For $DAX$, neither test rejects uniformity at the $5\%$ significance
level (Table 12). For $SPX$, both tests reject normality, suggesting the
data may not follow a uniform distribution (Table 12). Despite failing
the tests, we decided to proceed regardless as uniformity tests with
large samples can reject for small deviations that have limited impact
on $VaR$; many copula workflows proceed if diagnostics still broadly
look uniform and dependence fitting is stable[@blinov2014].

A bivariate copula is fitted via maximum likelihood on the
pseudo-observations, with model selection via AIC, Kendall's $\tau$, and
tail dependence coefficients. The candidate copulas considered span over
a broad set of families. Some common examples include:

-   Gaussian(implicit): elliptical, no tail dependence

-   Student-$t$ (implicit): elliptical, symmetrical tail dependence

-   Clayton (explicit): lower-tail dependence

-   Gumbel (explicit): upper-tail dependence

-   Frank (explicit): no tail dependence, flexible central dependence

-   Survival BB1(implicit): asymmetric non-linear dependence in tails
    (two-parameter "Clayton-- Gumbel" family, rotated via survival)

Historical $VaR$ is included as a nonparametric benchmark, though being
unconditional and backward-looking, it is regime-dependent and may be
unreliable as a predictor in the presence of volatility clustering and
structural change.

```{r, results='hide'}
model <- BiCopSelect(uspx, udax,
                     familyset = NA,
                     selectioncrit = "AIC",
                     indeptest = TRUE,
                     level = 0.05)

model
#Simulate N draws from the selected copula

N <- 1350
set.seed(1)

usim <- BiCopSim(N,
                  family = model$family,
                  par    = model$par,
                  par2   = model$par2)


```

The selected copula is Survival BB1 structure, implying asymmetrical tail dependence:

$$
\hat{C} = C_{\mathrm{sBB1}}\!\left(u_1,u_2;\,\hat{\theta},\hat{\delta}\right),
\qquad
\hat{\theta}=0.24,\ \hat{\delta}=1.85,
\qquad
\hat{\tau}=0.52.
$$

As a qualitative validation, $N=1350$ pairs are simulated from $\hat{C}$
and compared visually against the observed PIT pairs to confirm that the
fitted copula reproduces the empirical dependence pattern (Figure 7).

```{r,fig.width=10, fig.height=5, echo=FALSE, fig.cap= "Scatterplots of the original pseudo-observations and pseudo-observations simulated from the fitted Survival BB1 copula. The simulated sample visually matches the empirical dependence structure, supporting the adequacy of the selected copula model."}
# simulated uniforms should show the same dependence pattern
par(mfrow = c(1,2))
plot(data.frame(uspx,udax), pch=16, cex=0.35,
     main="Original pseudo-observations", xlab="uSPX", ylab="uDAX")
plot(data.frame(usim[,1],usim[,2]), pch=16, cex=0.35,
     main="Simulated from fitted copula", xlab="u1sim", ylab="u2sim")
```

## 2.8 VaR estimation

Let $\bigl(U^{(i)}_{\mathrm{SPX}},\,U^{(i)}_{\mathrm{DAX}}\bigr)$ be
$N_{sim} =50,000$ simulated draws from the fitted copula $\hat{C}$, for
$i=1,2,\ldots,N_{sim}$ Monte Carlo scenarios. Each draw is mapped back
to the innovation scale via the inverse fitted marginal CDF,

$$
Z^{(i)}_{j} = \hat{F}^{-1}_{Z_j}\!\left(U^{(i)}_{j}\right),
\qquad j\in\{\mathrm{SPX},\mathrm{DAX}\},
$$

the one-step-ahead return for scenario $i$ is reconstructed as

$$
r^{(i)}_{j,T+1} = \hat{\mu}_{j,T+1} + \hat{\sigma}_{j,T+1}\,Z^{(i)}_{j},
$$

Where $\hat{\mu}_{j,T+1}$ and $\hat{\sigma}_{j,T+1}$ are the ARMA--GARCH
predicted conditional mean and volatility $T+1$. Portfolio returns are
then formed as an equal-weight combination of
$r^{(i)}_{\mathrm{SPX},T+1}$ and $r^{(i)}_{\mathrm{DAX},T+1}$, and $VaR$
is estimated as the corresponding empirical quantile.

To assess the sensitivity to the choice of dependence structure, $VaR$
estimates are computed under the Survival BB1 copula and five
alternatives (Gaussian, Student-$t$, Clayton, Gumbel, and Frank),
alongside a historical empirical benchmark.

```{r, results='hide'}
set.seed(1)
Nsim <- 50000

# last-step mu/sigma (one-week ahead demo)
mu1  <- as.numeric(tail(fitted(model13), 1)); sig1 <- as.numeric(tail(volatility(model13), 1))
mu2  <- as.numeric(tail(fitted(model23), 1)); sig2 <- as.numeric(tail(volatility(model23), 1))

# fit common copulas on your PIT data (reuses uspx, udax)
fit_gauss   <- BiCopEst(uspx, udax, family=1)
fit_t       <- BiCopEst(uspx, udax, family=2)
fit_clayton <- BiCopEst(uspx, udax, family=3)
fit_gumbel  <- BiCopEst(uspx, udax, family=4)
fit_frank   <- BiCopEst(uspx, udax, family=5)

# simulate U for each copula
U_sel     <- BiCopSim(Nsim, family=model$family, par=model$par, par2=model$par2)
U_gauss   <- BiCopSim(Nsim, family=1, par=fit_gauss$par)
U_t       <- BiCopSim(Nsim, family=2, par=fit_t$par, par2=fit_t$par2)
U_clayton <- BiCopSim(Nsim, family=3, par=fit_clayton$par)
U_gumbel  <- BiCopSim(Nsim, family=4, par=fit_gumbel$par)
U_frank   <- BiCopSim(Nsim, family=5, par=fit_frank$par)

# VaR from U sample
VaR_from_U <- function(U){
  z1 <- qsstd(U[,1], nu=nuspx, xi=xispx)
  z2 <- qsstd(U[,2], nu=nudax, xi=xidax)
  r1 <- mu1 + sig1*z1
  r2 <- mu2 + sig2*z2
  rP <- log(1+0.5*(exp(r1)-1) + 0.5*(exp(r2)-1))
  quantile(rP, c(0.01, 0.05))
}

VaR_sel     <- VaR_from_U(U_sel)
VaR_gauss   <- VaR_from_U(U_gauss)
VaR_t       <- VaR_from_U(U_t)
VaR_clayton <- VaR_from_U(U_clayton)
VaR_gumbel  <- VaR_from_U(U_gumbel)
VaR_frank   <- VaR_from_U(U_frank)

VaR_hist <- quantile(returns$rP_rebal, c(0.01, 0.05), na.rm=TRUE)

table <- rbind(
  setNames(as.data.frame(t(VaR_sel)),     c("1%","5%")),
  setNames(as.data.frame(t(VaR_gauss)),   c("1%","5%")),
  setNames(as.data.frame(t(VaR_t)),       c("1%","5%")),
  setNames(as.data.frame(t(VaR_clayton)), c("1%","5%")),
  setNames(as.data.frame(t(VaR_gumbel)),  c("1%","5%")),
  setNames(as.data.frame(t(VaR_frank)),   c("1%","5%")),
  setNames(as.data.frame(t(VaR_hist)),    c("1%","5%"))
)

row.names(table) <- c(paste0("Selected: ", model$familyname),
                    "Gaussian","Student_t","Clayton","Gumbel","Frank","Historical")
table
```

\begin{table}[h!]
\centering
\begin{tabular}{lcc}
\toprule
Copula & VaR$_{0.01}$ & VaR$_{0.05}$ \\
\midrule
Selected: Survival BB1 & -0.04979521 & -0.02890508 \\
Gaussian               & -0.04829552 & -0.02886399 \\
Student-\textit{t}     & -0.04937962 & -0.02863751 \\
Clayton                & -0.05124481 & -0.02983666 \\
Gumbel                 & -0.04421892 & -0.02757014 \\
Frank                  & -0.04366806 & -0.02874826 \\
Historical             & -0.07400532 & -0.04120645 \\
\bottomrule
\end{tabular}
\caption{Value-at-Risk estimates at the 1\% and 5\% levels for the selected Survival BB1 copula, the alternative copula models, and the historical benchmark.}
\label{tab:copula_var}
\end{table}

Under the Survival BB1 copula, $VaR_{0.95}=0.0289$ and
$VaR_{0.99}=0.0498$ (Table 13). Alternative copulas yield broadly
consistent estimates; the Clayton copula produces the most conservative
model-based estimate $(Q_{0.01}=−0.0512)$, consistent with its lower
tail dependence, whilst the Frank and Gumbel copulas produce the least
conservative $(Q_{0.01}=−0.0437$ and $-0.0442$ respectively$)$.

The historical benchmark is substantially more conservative
$(VaR_{0.95}=0.0412, VaR_{0.99}=0.0740)$, as it pools realised returns
conditionally across the full 2000-2025 sample, including crisis
periods, rather than conditioning on the current volatility state
$\hat{\sigma}_{j,T+1}$ .

## 2.9 Model evaluation

Backtesting evaluates $VaR$ model adequacy by comparing predicted loss
thresholds against realised returns. The Kupiec test assesses whether
the observed exceedance rate is consistent with the nominal coverage
level, whilst the Christoffersen test additionally examines whether
exceedances are independently distributed over time, detecting clustered
violations that would indicate a failure to capture volatility dynamics.
$VaR$ forecasts at $α=0.05$ and $α=0.01$ are backtested over $T=1355$
weekly observations.

```{r, results='hide'}

set.seed(1)
Nsim <- 50000
U <- BiCopSim(Nsim, family=model$family, par=model$par, par2=model$par2)

# sstd shocks
p1 <- model13@fit$par; p2 <- model23@fit$par
z1 <- qsstd(U[,1], nu=unname(p1["shape"]), xi=unname(p1["skew"]))
z2 <- qsstd(U[,2], nu=unname(p2["shape"]), xi=unname(p2["skew"]))

# 1-step paths + realised portfolio return
mu1 <- as.numeric(fitted(model13)); s1 <- as.numeric(volatility(model13))
mu2 <- as.numeric(fitted(model23)); s2 <- as.numeric(volatility(model23))
Tn <- min(length(mu1), length(mu2), nrow(returns))

rP_real <- returns$rP_rebal[2:Tn]

# iterate over weeks calculating VaR
VaRret <- function(q){
  v <- numeric(Tn-1)
  for(t in 1:(Tn-1)){
    r1 <- mu1[t] + s1[t]*z1
    r2 <- mu2[t] + s2[t]*z2
    rP <- log(1 + 0.5*(exp(r1)-1) + 0.5*(exp(r2)-1))
    v[t] <- as.numeric(quantile(rP, probs=q))
  }
  v
}

VaR_95 <- VaRret(0.05)
VaR_99 <- VaRret(0.01)

vt95 <- VaRTest(alpha=0.05, actual=rP_real, VaR=VaR_95)
vt99 <- VaRTest(alpha=0.01, actual=rP_real, VaR=VaR_99)

data.frame(
  level=c("95%","99%"),
  T=c(length(rP_real), length(rP_real)),
  expected_exceed=c(vt95$expected.exceed, vt99$expected.exceed),
  observed_exceed=c(vt95$actual.exceed, vt99$actual.exceed),
  exceed_rate=c(vt95$actual.exceed/length(rP_real), vt99$actual.exceed/length(rP_real)),
  Kupiec_p=c(vt95$uc.LRp, vt99$uc.LRp),
  Christoffersen_p=c(vt95$cc.LRp, vt99$cc.LRp)
)
```

\begin{table}[h!]
\centering
\resizebox{\textwidth}{!}{%
\begin{tabular}{lcccccc}
\toprule
Level & $T$ & 
\begin{tabular}{c}Expected \\ Exceedances\end{tabular} &
\begin{tabular}{c}Observed \\ Exceedances\end{tabular} &
\begin{tabular}{c}Exceedance \\ Rate\end{tabular} &
\begin{tabular}{c}Kupiec \\ $p$-value\end{tabular} &
\begin{tabular}{c}Christoffersen \\ $p$-value\end{tabular} \\
\midrule
95\% & 1355 & 67 & 77 & 0.05682657 & 0.2587006 & 0.06020675 \\
99\% & 1355 & 13 & 16 & 0.01180812 & 0.5154702 & 0.32996970 \\
\bottomrule
\end{tabular}}
\caption{Backtesting of 1-day-ahead VaR forecasts.}
\label{tab:var_backtest}
\end{table}

\newpage
At the $95\%$ level, $77$ exceedances were observed against $67$
expected $(rate =0.0568)$, with the Kupiec and Christoffersen tests
returning $p$-values of $0.259$ and $0.060$ respectively, neither
rejecting at the $5\%$ level (Table 14). At the $99\%$ level, 16
exceedances were observed against 13 expected $(rate =0.0118)$, with
$p$-values of $0.515$ and $0.330$ respectively. At both confidence
levels, the model produces $VaR$ forecasts consistent with empirical
exceedance frequency and independence.

## 2.10 Limitations

Several limitations warrant acknowledgement. Uniformity diagnostics
indicate residual departure from $Uniform(0,1)$ for the $SPX$
pseudo-observations, suggesting marginal misspecification that may
affect copula estimation and tail inference. Model-based $VaR$ estimates
are also considerably less conservative than historical empirical
quantiles, reflecting the fact that the model conditions on current
GARCH volatility rather than the unconditional return distribution
across all regimes. Finally, copula selection relied on AIC and visual
diagnostics, no formal goodness-of-fit test was conducted, representing
a natural avenue for further refinement.

# 3. Conclusion

Using weekly $SPX$ and $DAX$ returns over $2000-2025$, ARMA--GARCH
marginal models with skewed Student-$t$ innovations were fitted to each
series, capturing volatility clustering, heavy tails, and serial
dependence. Following the PIT transformation, AIC selected a Survival
BB1 copula with $(\hat{\theta},\hat{δ})=(0.24,1.85)$ and Kendall's
$\tau = 0.52$, reflecting strong asymmetric tail dependence between the
two indices. Monte Carlo simulation $(N=50,000)$ yielded
$VaR_{0.95}=0.0289$ and $VaR_{0.99}=0.0498$ for an equal-weight
rebalanced portfolio, with Kupiec and Christoffersen backtests
confirming exceedance frequency and independence consistent with nominal
coverage at both levels. \newpage

# 4. References
